Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints

### 2020-01-27 2019/NUMTA_Barkalov/Special Issure/article.tex
figures are corrected

\begin{figure}[h]
	\center{\includegraphics[width=0.7\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ (a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ (b)}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ b)}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ b)}
	\end{minipage}
	\caption{Solving a problem of the “Simple” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_simple}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ b)}
	\end{minipage}
	\caption{Solving a problem of the “Hard” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_hard}
\end{figure}

	Also, the points of the trials performed by (a) the original index algorithm (IA) and by (b) the method with dual estimate of the Lipschitz constant (IA-DL) are marked in Figs. \ref{fig:examples_simple} and \ref{fig:examples_hard}. To solve the problems shown in Figs. \ref{fig:examples_simple} and \ref{fig:examples_hard}, the IA method required 643 and 1166 iterations whereas the IA-DL method solved similar problems for 373 and 712 iterations, respectively.

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ b)}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 3$}
	\label{fig:operational_characteristics_N=3}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ b)}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 5$}
	\label{fig:operational_characteristics_N=5}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ a)}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ b)}
	\end{minipage}
	\caption{Frequency of the use of different estimates of the Lipschitz constant for the problems of the “Simple” class (a) and of the “Hard” class (b) for $N=4$}
	\label{fig:frequency_N=4}
\end{figure}

>>>>

\begin{figure}%[h]
	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}%[h]
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Simple” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_simple}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Hard” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_hard}
\end{figure}

	Also, the points of the trials performed by (a) the original index algorithm (IA) and by (b) the method with dual estimate of the Lipschitz constant (IA-DL) are marked in Figs. \ref{fig:examples_simple} and \ref{fig:examples_hard}. To solve the problems shown in Figs. \ref{fig:examples_simple} and \ref{fig:examples_hard}, the IA method required 643 and 1132 iterations whereas the IA-DL method solved similar problems for 373 and 443 iterations, respectively.

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 3$}
	\label{fig:operational_characteristics_N=3}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_5} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 5$}
	\label{fig:operational_characteristics_N=5}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Decision_rule_gklssC_4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Decision_rule_gklshC_4} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Frequency of the use of different estimates of the Lipschitz constant for the problems of the “Simple” class (a) and of the “Hard” class (b) for $N=4$}
	\label{fig:frequency_N=4}
\end{figure}
